% Appendix C — soft-mode escape methodology (H3As Im-3m -> R3m).
\section{Appendix C --- The soft-mode-escape methodology}
\label{app:C}

This appendix documents the campaign's most transferable contribution: a
procedure for converting a dynamically \emph{unstable} high-symmetry
candidate into a \emph{stable} lower-symmetry polymorph that retains strong
coupling. The worked example is \ce{H3As}.

\paragraph{The instability.}
\ce{H3As} in the high-symmetry $Im\bar{3}m$ structure is a dynamical saddle
point: a $4^3$-$q$ \texttt{ph.x} census finds imaginary modes ($25/192$
imaginary, minimum $-\SI{1267}{\per\centi\meter}$). A naive Allen--Dynes
read of the unconverged $Im\bar{3}m$ numbers gives only $\lambda \approx
0.97$, $T_c \approx \SI{30}{\kelvin}$ --- the lattice is collapsing before
strong coupling sets in.

\paragraph{The escape (method).}
\begin{enumerate}\itemsep0.25em
\item Identify the softest (most negative) phonon eigenvector at the
  unstable $q$-point. This is the direction along which the high-symmetry
  cell wants to distort.
\item Displace the structure along that eigenvector and relax. The symmetry
  drops --- here $Im\bar{3}m \to R3m$ (rhombohedral), a trigonal distortion
  of the cubic cell.
\item Re-run the full electron--phonon calculation in the distorted cell. If
  the distortion has removed the imaginary modes, the polymorph is the true
  ground state at that pressure and its $(\lambda, \omega_{\log})$ are
  physical.
\end{enumerate}

\paragraph{The result.}
The $R3m$ \ce{H3As} polymorph is dynamically stable ($14/14$ $q$-points
clear) and carries genuine strong coupling: $\lambda = 1.65$, $\omega_{\log}
= \SI{450}{\kelvin}$. The Allen--Dynes critical temperature is verified to
libm precision (g5):

\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,frame=single,xleftmargin=0pt]
$ hexa verify --expr allen_dynes_tc 1.65 450 0.10 55.88 --tol 0.01
  calc   = 55.8806  ~= expected 55.88  (within tol 0.01, |D|=0.000567718)
  tier   = GREEN SUPPORTED-NUMERICAL  (H3As R3m soft-mode-escape polymorph)
\end{lstlisting}

\noindent The escape recovers a stable, strongly-coupled phase
($T_c = \SI{55.9}{\kelvin}$) from a candidate that the high-symmetry screen
would have discarded as ``unstable, weak.'' The coupling nearly doubled
($0.97 \to 1.65$) once the lattice was allowed to find its real ground state.

\paragraph{Why this matters beyond \ce{H3As}.}
High-throughput hydride screens routinely reject candidates on a single
high-symmetry phonon calculation. Soft-mode escape says that a negative
eigenvalue is an \emph{instruction}, not a verdict: it points to the
lower-symmetry polymorph that should be evaluated instead. The procedure is
generic --- it requires only the soft eigenvector, which \texttt{ph.x}
already produces --- and it is the methodological complement to the N5 wall
(Appendix~\ref{app:D}): the wall says strong harmonic coupling is often a
soft-mode artifact, and escape is the disciplined way to find out whether a
stable polymorph survives underneath.

\paragraph{Honest scope.}
\ce{H3As}-$R3m$ at \SI{56}{\kelvin} is not a record and not ambient. The
methodology, not the number, is the contribution. The $R3m$ verdict is a
campaign-internal DFT result pending an independent atlas-registration
record; the $T_c$ value itself is \tierGreen\ closed-form, but the structural
identification carries the usual DFT-level (not measured) status (R4).
